---
title: "Assesments-Wk2"
author: "P N H"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quiz 2

error reason: 'x' is a vector of length 10 and 'if' can only test a single logical statement.

```{r}
x <- 1:10
if(x > 5) {
        x <- 0
}
```

```{r}
f <- function(x) {
        g <- function(y) {
                y + z
        }
        z <- 4
        x + g(x)
}
z <- 10
f(3)
```
This code evaluates y=10 for any x>3 and y=NA for x<3
```{r}
x <- 5
y <- if(x < 3) {
        NA
} else {
        10
}
```


```{r}
h <- function(x, y = NULL, d = 3L) {
        z <- cbind(x, d)
        if(!is.null(y))
                z <- z + y
        else
                z <- z + f
        g <- x + y / z
        if(d == 3L)
                return(g)
        g <- g + 10
        g
}
```
Which symbol in the above function is a free variable? z is not a free variable. Why?

What type of scoping rule for resolving free variables are used in R? Lexial

---------------------------------------------------------------------------------
Assignment 1, Week 2, R Programming
https://www.coursera.org/learn/r-programming/supplement/amLgW/programming-assignment-1-instructions-air-pollution
Faster way at the end of this document , also inside a link in this doc: https://github.com/rdpeng/practice_assignment/blob/master/practice_assignment.rmd

part 3: https://www.coursera.org/learn/r-programming/discussions/weeks/2/threads/r5z9Go2UQQ2c_RqNlHENZA


*Assignment 1, Part1*
This is Correct
Write a function named 'pollutantmean' that calculates the mean of a pollutant (sulfate or nitrate) across a specified list of monitors. The function 'pollutantmean' takes three arguments: 'directory', 'pollutant', and 'id'. Given a vector monitor ID numbers, 'pollutantmean' reads that monitors' particulate matter data from the directory specified in the 'directory' argument and returns the mean of the pollutant across all of the monitors, ignoring any missing values coded as NA.


#visit: https://github.com/lgreski/datasciencectacontent/blob/master/markdown/rprog-breakingDownMakeVector.md
#The overall objective of the assignment 2 is to demonstrate the concept of lexical scoping.

#https://github.com/lgreski/datasciencectacontent/blob/master/markdown/rprog-discussPollutantmean.md

-----
WHY?

pollutantmean("specdata", "sulfate", 1:10)
why: 4.064  but my answer is 4.05555

PART 1: CORRECT
```{r}
library(fs)

pollutantmean <-function(directory, pollutant, id=1:332, removaNA=TRUE){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame()
        print(df)
        csv <- data.frame()

        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                #print(csv[,pollutant])
                df <-(rbind(df, csv, na.rm=TRUE))
                print( glue("finieshed j= {j}"))
        }

        print(tail(df))
        means<-(mean(df[,pollutant], na.rm=TRUE))
        print(means)
}
        library(glue)
#        glue("The overall mean value across the given range for sulfate is:  {means[2]}, and for nitrate is: {means[3]}.")

#source("pollutantmean.R") #orwhatever the file name is!!!
#pollutantmean (directory="~/dev/datasciencecoursera/data/specdata/", pollutant="nitrate", id=c(1,3))

```



*Assignment 1, Part2*
Write a function that reads a directory full of files and reports the number of completely observed cases in each data file. The function should return a data frame where the first column is the name of the file and the second column is the number of complete cases. 
2nd method: I could have also rbinded the data frames as done in part 1


```{r}
library(fs)

complete <-function(directory, pollutant, id){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame(matrix(ncol =2, nrow = length(id)))
        colnames(df) <- c("ID", "No Comp. Cases")
        row_num <- 1
        
        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                print(sum_existing <-sum(!is.na(csv[,pollutant])))

                row.names(df)[row_num] <- file_name
                df[row_num,1]<- j
                df[row_num,2]<- sum_existing
                row_num <- row_num +1
        }
                print(df)

}
#complete(directory="~/dev/datasciencecoursera/data/specdata/", "sulfate", id=c(1))
```




 ---------------------

Part 3
Write a function that takes a directory of data files and a threshold for complete cases and calculates the correlation between sulfate and nitrate for monitor locations where the number of completely observed cases (on all variables) is greater than the threshold. The function should return a vector of correlations for the monitors that meet the threshold requirement. If no monitors meet the threshold requirement, then the function should return a numeric vector of length 0. 

corr("specdata",500) should only produce correlations for the sensors with at least 501 complete cases of sulfate and nitrate.

The logic should go like this:

Read a sensor file

Determine whether it has sufficient complete observations to exceed the threshold, and if so, 

calculate a correlation between nitrate and sulfate

save the output to a vector that you can return to the parent environment once all sensor files have been processed.

You can source complete.R with the source() function. One way to use complete() within corr() would be with a function that follows this outline:

run complete() and save the result to a data frame

extract the sensor IDs exceeding the threshold value from the saved data frame

calculate correlations for each sensor in the id vector created in step 2, build them into a vector, and return to the parent environment. correlate the nitrate and sulfate values for only those monitors with *complete cases greater than the threshold, including NA values*. If you wipe out the NA values, your nitrate and sulfate vectors are going to end up being different lengths, so the cor() function will throw an error.Instead, in the cor() function set the "use" parameter equal to "complete.obs", which tells cor() to simply disregard NA values by casewise deletion. So, cor() will calculate the correlation of only complete cases of that monitor.You want to utilize the  use = "complete.obs" option in the cor function, otherwise you will just get lots of NA entries. 


The cor function applied to 2 vectors will use all the entries of the two vectors including any NA's unless you "tell it" to do otherwise with the use argument. use = "na.or.complete" is an appropriate choice here.

Correct
length is number of coloumns.
In the asignment the new data is added to a new column and not a row
use round(cr, 3) to round the result to 3 decimal places

RNGversion("3.5.1")
```{r}
corr <- function(directory, threashold=0, id=1:332){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame(matrix(ncol =3, nrow = length(id)))
        colnames(df) <- c("ID", "S Cases", "Correlation")
        row_num <- 1
        
        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                sum_existing <-sum(!is.na(csv[,"sulfate"]))

                if  (sum_existing>threashold){
                        row.names(df)[row_num] <- file_name
                        df[row_num,1]<- j
                        df[row_num,2]<- sum_existing
                        df[row_num,3]<- cor(csv[,"sulfate"], csv[,"nitrate"],  use = "na.or.complete")
                        row_num <- row_num +1
                }
        }
        print(df)    
}
#corr(directory="~/dev/datasciencecoursera/data/specdata/", threashold=100, id=c(1))
```

***************************************************************************************

Part 1
This method is WRONG as the average of average is not the same as average!

```{r}
#Choosing few files

library(magrittr) # %>%
library(fs) #path

#colummean <-function(y,id, removeNA=TRUE){ #y is a dataframe or a matrix
id <- as.integer(c(1,4))
dir_path <- path("~/dev/datasciencecoursera/data/specdata/")
list_csv_name <-list.files(dir_path, pattern="csv")

df <- data.frame(matrix(ncol =4, nrow = length(id)))
colnames(df) <- c("Date", "sulfate", "nitrate", "ID")
row_num <- 1

for (j in id){
        print(j)
        file_name <- list_csv_name[j]
        print(file_name)
        csv<- read.csv(dir_path/file_name)
        # b <- as.integer(unlist(strsplit(x, ".", fixed=TRUE)))
        nc <-ncol(csv) #number of columns
        means <-numeric(nc) #initialize the loop. An empty numeric vector.To store the means for each coloumn
        
        for(i in 1:nc){ #for each column do the following
                means[i] <-(mean(csv[,i], na.rm=TRUE))
                print(means[i])
        }
        
        print(means)
        row.names(df)[row_num] <- file_name
        df[row_num, ]<- means
        row_num <- row_num +1 
        }
        
         for(i in 1:nc){ #for each column do the following
                means[i] <-(mean(df[,i], na.rm=TRUE))
                print(means[i])
        }
library(glue)
glue("The overall mean value across the given range for sulfate is:  {means[2]}, and for nitrate is: {means[3]}.")

```



PART 1 
Making the above into functions
This method is Wrong as the average of average is not the same as average!

```{r}
library(fs)

pollutantmean <-function(directory, pollutant, id, removaNA=TRUE){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame(matrix(ncol =4, nrow = length(id)))
        colnames(df) <- c("Date", "sulfate", "nitrate", "ID")
        row_num <- 1
        
        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                # print(csv[,pollutant][303])
                
                means <-(mean( csv[,pollutant], na.rm=TRUE))
                
                row.names(df)[row_num] <- file_name
                df[row_num, ]<- means
                row_num <- row_num +1
                print(df)
        }
 
        means<-(mean(df[,pollutant], na.rm=TRUE))
        print(means)

        
}
        library(glue)
        glue("The overall mean value across the given range for sulfate is:  {means[2]}, and for nitrate is: {means[3]}.")

#pollutantmean (directory="~/dev/datasciencecoursera/data/specdata/", pollutant="nitrate", id=1)

```



part 2: This is counting total number of complete cases for both pollutants of sulfate and nitrate required for part 3
```{r}
rm(list=ls())
library(fs)

complete2 <-function(directory, pollutant1, pollutant2=0, id){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame(matrix(ncol =4, nrow = length(id)))
        colnames(df) <- c("ID", "Pol.1","Pol2", "Pol_tot")
        row_num <- 1
        
        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                sum_existing1 <-sum(!is.na(csv[,pollutant1]))
                sum_existing2 <-sum(!is.na(csv[,pollutant2]))
                sum_existing <-sum(sum_existing1,sum_existing2)
                
                row.names(df)[row_num] <- file_name
                df[row_num,1]<- j
                df[row_num,2]<- sum_existing1
                df[row_num,3]<- sum_existing2
                df[row_num,4]<- sum_existing
                row_num <- row_num +1
        }
        print(df)
        
}
#complete2(directory="~/dev/datasciencecoursera/data/specdata/", "sulfate", id=c(1))
```

Part 3
This is incorrect. It has extra steps that are not required in the assignment. The sample only cares about sulfate numbers of total NA cases.
```{r}
corr <- function(directory, threashold=0, id=1:332){
        
        list_csv_name <-list.files(path=directory, pattern="csv")
        df <- data.frame(matrix(ncol =5, nrow = length(id)))
        colnames(df) <- c("ID", "Pol.1","Pol2", "Pol_tot", "Correlation")
        row_num <- 1
        
        for (j in id){
                file_name <- list_csv_name[j]
                file <- paste(directory,file_name,sep="/")
                csv<- read.csv(file)
                sum_existing1 <-sum(!is.na(csv[,"sulfate"]))
                sum_existing2 <-sum(!is.na(csv[,"nitrate"]))
                sum_existing <-sum(sum_existing1,sum_existing2)
                
                if  (sum_existing>threashold){
                        row.names(df)[row_num] <- file_name
                        df[row_num,1]<- j
                        df[row_num,2]<- sum_existing1
                        df[row_num,3]<- sum_existing2
                        df[row_num,4]<- sum_existing
                        df[row_num,5]<- cor(csv[,"sulfate"], csv[,"nitrate"],  use = "na.or.complete")
                        row_num <- row_num +1
                }
                
        }
        print(df)    
}
#corr(directory="~/dev/datasciencecoursera/data/specdata/", threashold=100, id=c(1))
```
