---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Principlace of EDA: Quick and dirty
first to show some comparison, 
the second was to show a mechanism, 
Third

We DON'T use exploratory graphs to communicate results.

 one-dimensional graphs that summarize data. 
 Two dimensional graphs include scatterplots
 Some graphs have more than two-dimensions. These include overlayed or multiple two-dimensional plots and spinning plots.
```{r}
head(pollution)
#        pm25  fips region longitude latitude
# 1  9.771185 01003   east -87.74826 30.59278
# 2  9.993817 01027   east -85.84286 33.26581
# 3 10.688618 01033   east -87.72596 34.73148
# 4 11.337424 01049   east -85.79892 34.45913
# 5 12.119764 01055   east -86.03212 34.01860
# 6 10.827805 01069   east -85.35039 31.18973

names(pollution)# gives col names
summary(pollution)
summary(pollution$pm25) # summary of one variable only
ppm # To save you a lot of typing we've saved off pollution$pm25 in this variable
quantile(ppm) #quantile gives the quartiles

boxplot(ppm, col="blue") # show quartile data that summary and quantile did. The lower and upper edges of the blue box respectively show the values of the 25% and 75%, middle hrizontal line=median or 50%. the box height is the difference between the 75th and 25th quantiles.
abline(h = 12) #OVERLAY ABILITY: adds a horizontal line to the plot. 12 is the EPA standard for air quality

hist(ppm, col="green") #gives the most frequent pollution counts
rug(ppm) #how many data points are in each bucket and where they lie through densityo f tick marks
hist(ppm, col="green", breaks =100) #the number of buckets to split the data into
abline(v=12, lwd=2) #adds a vertical line at 12 and width 2. not very visiable
abline(v=median(ppm), col="magenta", lwd=4) # run the hist() . then vertical line at med, and pink coour


reg <- table(pollution$region) #more pollution in the east
# east west 
#  442  134 
barplot(reg, col="wheat", main="Number of Counties in Each Region")

boxplot(pollution$pm25~pollution$region, col="red")
boxplot(pm25~region, data=pollution, col="red") #same as

#multiple plots:
par(mfrow=c(2,1),mar=c(4,4,2,1)) #set up the plot window for two rows and one column, margins
east <- subset(pollution,region=="east")  #"returns subsets of vectors, matrices or data frames which meet conditions", only looks for boolean of region= east.

#        pm25  fips region longitude latitude
# 1  9.771185 01003   east -87.74826 30.59278
# 2  9.993817 01027   east -85.84286 33.26581
west <- subset(pollution,region=="west")

hist(east$pm25, col = "green") #The command par told R we were going to have one column with 2 rows, so it placed this histogram in the top position
hist(subset(pollution,region=="west")$pm25, col = "green") #same as, but labels are different
 
#scatter plot
with(pollution, plot(latitude, pm25)) #aregments of atitude, pm2 are in pollution dataframe
abline(h=12, lwd=2, lty=2)  #adds line, width and type
plot(pollution$latitude, ppm, col = pollution$region) #distinguish between counties in the east and those in the west. the high (greater than 50) and low (less than 25) latitudes are both red (west)
abline(h=12, lwd=2, lty=2)
#two scatter plots
par(mfrow = c(1, 2), mar = c(5, 4, 2, 1))
plot(west$latitude,west$pm25, main="West" ) #First Plot
plot(east$latitude,east$pm25, main="East" )
```
3. Graphics Devices in R 
```{r}
windows() #call screen device
```